Kaluza--Klein Cosmology from five-dimensional Lovelock--Cartan Theory

### 2016-09-14 KKosmology.tex
Bouncing solutions revised.

@@

For $u_\pm>0$ (see Fig.~\ref{cond_u_pos}) the solutions are periodic. In that case, the scale factor $a(t)$ starts from a singular point and reaches a future one after a time $t_{\rm bounce}=\pi/\sqrt{u_\pm}$. Depending on the particular region on the parameter space, the scale factor undergoes an expanding and contracting age, allowing to reach a minimum value without collapsing, to expand again before a big crunch. Otherwise, it reaches a maximum value before its collapse. It also shows an inflationary phase at early times when the scale factor grows exponentially. The bouncing behaviour of the extra dimension is given by the scalar field in a phase difference of $\pi/2$ with respect to $a(t)$.

\begin{figure}[H]
  \includegraphics[width=\linewidth]{cond_u_pos.pdf}
  \caption{The figure on the left denotes the region on the parameter space where $u_+>0$, while the figure on the right shows where $u_->0$. We have normalized $\alpha_1 = 1$.}
  \label{cond_u_pos}
\end{figure}
\begin{figure}[H]
  \includegraphics[width=\linewidth]{a_u_pos.pdf}
  \caption{blablabla}
  \label{a_u_pos}
\end{figure}
\begin{figure}[H]
  \includegraphics[scale=.45]{phi_u_pos.pdf}
  \caption{Behavior of the $\phi$ field as a function of the scaled time, for $u_\pm > 0$.}
  \label{phi_u_pos}
\end{figure}

@@

For $u_\pm\leq 0$ (see Fig.~\ref{cond_u_neg}), depending on the sign of $\alpha_2$ and the $\pm$-sign in $u_\pm$, the scale factor will infinitely expand while the scalar field reaches an asymtotic vanishing value for late cosmic times, after a period of expansion. This is interpreted as the late time behaviour of the radius of the compact manifold $S^1$, which for this family of solutions provides a natural explanation for its scale. Otherwise, in contrast to the previous behaviour, the scale factor contracts to an asymtotic vanishing value while the scalar field has no upper bound. In the case of strictly negative $u_\pm$, for either contracting or expanding scale factors, their initial values remain finite and the solutions do not present an initial singularity. It has been reported that the Gauss--Bonnet term can prevent the Universe to expand from an initial singularity \cite{Deruelle:1986iv,Henriques:1986jw,Ishihara:1986if}, which is also the present case.

\begin{figure}[H]
  \includegraphics[scale=.45]{cond_u_neg.pdf}
  \caption{The figure on the left denotes the region on the parameter space where $u_+<0$, while the figure on the right shows where $u_-<0$. We have normalized $\alpha_1 = 1$.}
  \label{cond_u_neg}  
\end{figure}
\begin{figure}[H]
  \includegraphics[scale=.45]{a_u_neg.pdf}
  \caption{}
  \label{a_u_neg}
\end{figure}
\begin{figure}[H]
  \includegraphics[scale=.45]{phi_u_neg.pdf}
  \caption{}
  \label{phi_u_neg}
\end{figure}

\begin{figure}[H]
  \includegraphics[scale=.45]{a_u-_0.pdf} 
  \caption{Behaviour for $u_- = 0$.}
  \label{a_u_0}
\end{figure}
\begin{figure}[H]
  \includegraphics[scale=.45]{phi_u-_0.pdf}
  \caption{Behaviour for $u_- = 0$.}
  \label{phi_u_0}
\end{figure}

>>>>

Hereon, we will consider $\alpha_1$ to be positive,\footnote{In all the plots, we fixed $\alpha_1 = 1$.} because this parameter admits the interpretation of the Newton's gravitational constant.

@@

In Fig.~\ref{cond_u_pos} we show the allowed region, in the $\alpha_0$--$\alpha_2$ parameter space, for $u_\pm$ to be positive. 
\begin{figure}[H]
  \includegraphics[width=\linewidth]{cond_u_pos.pdf}
  \caption{Allowed regions, in the $\alpha_0$--$\alpha_2$ plane, for $u_+$ and $u_-$ to be positive defined. We have set $\alpha_1 = 1$.}
  \label{cond_u_pos}
\end{figure}

All the solutions for $u_\pm>0$ are periodic, as shown in the behavior of the scale factor in Fig.~\ref{a_u_pos}. In that case, the scale factor $a(t)$ starts from a singular point and reaches a future one after a time $t_{\rm bounce}=\pi/\sqrt{u_\pm}$.
Depending on the particular region on the parameter space, the scale factor undergoes an expanding and contracting age, allowing a small bounce without collapsing, before a big crunch. This is the case of the third curve in Fig.~\ref{a_u_pos}.
Otherwise, it expands and collapses in a \emph{simple} oscillatory way. One can also notice from the plot, that solutions parametrized by $u_+$ do not present inflationary phase, while those parametrized by $u_-$ do---the scale factor grows exponentially for a certain interval of time. Interestingly, the solutions with inflationary phases collapse in a damped form. Moreover, the bouncing behavior of the extra dimension is given by the scalar field in a phase difference of $\pi/2$ with respect to $a(t)$, as shown in Fig.~\ref{phi_u_pos}. It is worth to point that the \emph{small bounce} present in the scale factor is much prominent in the modulation of the extra dimension.
\begin{figure}[H]
  \includegraphics[width=\linewidth]{a_u_pos.pdf}
  \caption{Behavior of the scale factor, \(a(t)\), as a function of the scaled time, $\tau$, for $u_\pm > 0$.}
  \label{a_u_pos}
\end{figure}
\begin{figure}[H]
  \includegraphics[scale=.45]{phi_u_pos.pdf}
  \caption{Behavior of the $\phi$ field as a function of the scaled time, $\tau$, for $u_\pm > 0$.}
  \label{phi_u_pos}
\end{figure}

@@

None of the other cases, $u_- = 0$ and $u_\pm < 0$, present bouncing behavior.
The allowed moduli space for $u_+$ and $u_-$ to be negative is shown in Fig.~\ref{cond_u_neg}, while the restriction $u_- = 0$ sets $\alpha_2 = \tfrac{1}{12 \alpha_0}$.
Nonetheless, depending of the values of the parameters there are three distinctive conduct: (i)~eternal expansion, (ii)~eternal contraction, and (iii)~initial expansion with a final collapse.

The case $u_- = 0$ has two behaviors depending on whether $\alpha_0$ is positive or negative. As shown in Fig.~\ref{a_u_0}, for positive $\alpha_0$ the solutions fit into the third category above, while for negative $\alpha_0$ the solutions expand eternally. Notice that in the former case the Universe's fate is an asymptotic collapse, whilst in the later expands forever but shows an interesting evolution of the Hubble constant.
\begin{figure}[H]
  \includegraphics[scale=.45]{a_u-_0.pdf} 
  \caption{Behaviour for $u_- = 0$.}
  \label{a_u_0}
\end{figure}
\begin{figure}[H]
  \includegraphics[scale=.45]{phi_u-_0.pdf}
  \caption{Behaviour for $u_- = 0$.}
  \label{phi_u_0}
\end{figure}

For $u_\pm\leq 0$ (see Fig.~\ref{cond_u_neg}), depending on the sign
of $\alpha_2$ and the $\pm$-sign in $u_\pm$, the scale factor will
infinitely expand while the scalar field reaches an asymtotic
vanishing value for late cosmic times, after a period of
expansion. This is interpreted as the late time behaviour of the
radius of the compact manifold $S^1$, which for this family of
solutions provides a natural explanation for its scale. Otherwise, in
contrast to the previous behaviour, the scale factor contracts to an
asymtotic vanishing value while the scalar field has no upper
bound. In the case of strictly negative $u_\pm$, for either
contracting or expanding scale factors, their initial values remain
finite and the solutions do not present an initial singularity. It has
been reported that the Gauss--Bonnet term can prevent the Universe to
expand from an initial
singularity~\cite{Deruelle:1986iv,Henriques:1986jw,Ishihara:1986if},
which is also the present case.

\begin{figure}[H]
  \includegraphics[scale=.45]{cond_u_neg.pdf}
  \caption{Allowed regions, in the $\alpha_0$--$\alpha_2$ plane, for $u_+$ and $u_-$ to be positive defined. We have normalized $\alpha_1 = 1$.}
  \label{cond_u_neg}  
\end{figure}
\begin{figure}[H]
  \includegraphics[scale=.45]{a_u_neg.pdf}
  \caption{}
  \label{a_u_neg}
\end{figure}
\begin{figure}[H]
  \includegraphics[scale=.45]{phi_u_neg.pdf}
  \caption{}
  \label{phi_u_neg}
\end{figure}
